The Skew-Normal Approximation of the Binomial Distribution

### 2011-01-25 project/paper/method-of-moments.tex
Proofreading.
Tweaks to method of moments section.

From these, we can find $E(B^2)$: Recall that $Var(B) = E(B^2) - [E(B)]^2$ and
thus

@@

We now set the left hand side \eqref{eq:binomial-third-moment-lhs} and right
hand side \eqref{eq:binomial-third-moment-rhs} equal to each other and solve
for $E(B^3)$:

@@

The above equation \eqref{eq:solving-for-lambda} is a rational expression in
$\lambda^2$ that can be solved with either a considerable amount of manual
labor or, more efficiently, with a computer algebra system. Once we have
$\lambda^2$, then $\lambda$ is simply either the positive or negative square
root, as determined by the sign of $(1-2p)$. The sign can be explained with a
little assistance from Property \ref{prop:3}: When $p \to 0$, the binomial
skews left and converges toward the positive half normal, which by
\eqref{eq:p2-positive-half-normal} corresponds to a positive $\lambda$. When $p
\to 1$, the binomial skews right and converges toward the negative half normal,
which by \eqref{eq:p2-negative-half-normal} corresponds to a negative
$\lambda$. When $p = 0.5$, the binomial is symmetric and $\lambda$ is 0,
eliminating the need for a sign. Thus:

@@

One notable case where the above skew-normal approximation fails is when $p =
0.5$; the right side of \eqref{eq:solving-for-lambda} becomes undefined, and we
are unable to obtain $\lambda$. Fortunately, here we can fall back on our
intuition, which tells us that since the binomial is perfectly symmetrical, the
skew should be 0. A few observations support this conclusion: When $p = 0.5$,
equation \eqref{eq:lambda-solved} fails to yield a sign. Furthermore, when
$\lambda = 0$, equations \eqref{eq:sigma-solved} and \eqref{eq:mu-solved}
return us to the normal approximation ($\mu = np$ and $\sigma =
\sqrt{np(1-p)}$, respectively), which is after all a natural choice for a
symmetric binomial curve.

@@

Unfortunately, although better than the normal approximation, the skew-normal
is also not universally applicable. To obtain an estimate for $\lambda$, we
must put a few restrictions on $n$ and $p$.

>>>>

From these facts, we can find $E(B^2)$: Recall that $Var(B) = E[(B - E(B))^2] =
E(B^2) - [E(B)]^2$ and thus

@@

Since \eqref{eq:binomial-third-moment-rhs} and
\eqref{eq:binomial-third-moment-lhs} are equal, we can solve for $E(B^3)$:

@@

The above equation \eqref{eq:solving-for-lambda} is a rational expression in
$\lambda^2$ that can be solved with either a considerable amount of manual
labor or, more efficiently, with a computer algebra system. Once we have
$\lambda^2$, then $\lambda$ is simply the positive or negative square root, as
determined by the sign of $(1-2p)$.

The sign can be explained with a little assistance from Property \ref{prop:3}:
When $p \to 0$, the binomial shifts its weight towards the smaller values on
the left and converges toward the positive half normal, which by
\eqref{eq:p2-positive-half-normal} corresponds to a positive $\lambda$. When $p
\to 1$, the binomial shifts its weight towards the the larger values on the
right and converges toward the negative half normal, which by
\eqref{eq:p2-negative-half-normal} corresponds to a negative $\lambda$. When $p
= 0.5$, the binomial is symmetric and $\lambda$ is 0, eliminating the need for
a sign. Thus:

@@

One notable case where the above skew-normal approximation fails is when $p =
0.5$; the right side of \eqref{eq:solving-for-lambda} becomes undefined, and we
are unable to obtain $\lambda$. Fortunately, here we can fall back on our
intuition, which tells us that since the binomial is perfectly symmetrical, the
skew should be 0. A few observations support this conclusion: When $p = 0.5$,
equation \eqref{eq:lambda-solved} fails to yield a sign. Furthermore, when
$\lambda = 0$, equations \eqref{eq:sigma-solved} and \eqref{eq:mu-solved}
return us to the normal approximation ($\mu = np$ and $\sigma =
\sqrt{np(1-p)}$, respectively), which is after all a natural choice for a
symmetric binomial distribution.

@@

Unfortunately, the skew-normal is also not universally applicable. To obtain an
estimate for $\lambda$, we must put a few restrictions on $n$ and $p$.
